---
title: "Learning-R"
author: "Anil Akyildirim"
date: "7/16/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

The purpose of this document is to review the hands on lab quiz and practice R exercices in Week-1. 

## Questions

Question 1: Create a vector that contains 20 numbers. 

```{r}

number_vector <- c (1,2,3,4,5,1,2,3,4,5,6,7,8,9,10,11,12,13,13,12)
length(number_vector)
number_vector
```

Question 2: Use R to convert vector from question 1 into a character vector.


```{r}
number_vector <- as.character(number_vector)
typeof(number_vector)

```

Question 3: Use R to convert vector from question 1 into a vector of factors:
```{r}

number_vector <- as.factor(number_vector)
typeof(number_vector)
class(number_vector)
```
Question 4: Use R to show how many levels the vector in the previous has.

```{r}
nlevels(number_vector)
```
Question 5:  Use R to create a vector that takes vector from question 1 and performs on it the formula 3x2 -4x+1

Please note: In this case i realized i already converted the vector in question 1 into a factor type so i will create another numeric vector to calculate the equation.

```{r}
new_vector <- c(1:20)
new_vector

new_vector_2 <- 3*new_vector^2-4*new_vector+1
new_vector_2
```
Question 6: Create a named list that has several elements that are each able to be referenced by name:

```{r}

new_list <- list(title=c('Toy Story', 'Star Wars', 'Guardians of the Galaxy', 'Spiderman'), ratings=c(3,4,3,2))
new_list
```

Question 7: Create a dataframe with 4 columns - one each character, 3 factor levels, numeric and date. with 10 rows.

```{r}

title <- as.character(c('Toy Story', 'Star Wars', 'Guardians of the Galaxy', 'Spiderman', 'Incredibles 2',
           'Venom', 'Black Panther', 'Avengers', 'Iron Man', 'Avatar'))
release <- as.Date(c('2008-1-1', '2019-1-1', '2018-1-1', '2014-1-1', '1982-1-1',
                   '2007-1-1', '2015-1-1', '2016-1-1', '2019-1-1', '2017-1-1'))
genre <- c('comedy', 'adventure', 'action', 'comedy', 'adventure',
           'action', 'comedy', 'adventure', 'action', 'comedy')
ratings <- as.numeric(c(1,2,3,4,5,6,7,8,9,10))

df <- data.frame(title, release, genre, ratings)
head(df)
str(df)

# i see title has 10 factors. i need to make that a character.

df$title <- as.character(df$title)
str(df)
```
Question 8: Illustrate how to add a row with a value for the factor column that isnt already in the list. 

```{r}
new_movie <- data.frame(title='Batman', release='1996-1-1', genre='drama', ratings=10)
new_movie

new_movie$title <- as.character(new_movie$title)
new_movie$release <- as.Date(new_movie$release)

df <- rbind(df, new_movie)
df
```
Question 9: Show the code that would read in a CSV file called temperatures.csv 

Please note; there will not be an output of the CSV file as i do not have the csv in my local. 

``` {r}
df_new <- read.csv('temperatures', header=TRUE, sep=',')
df_new


```
Question 10: Use a loop to calculate the final balance, rounded to the nearest cent, in an account that earns 3.24% interest compounded monthly after six years if the original balance is $1500.
*Compount interes every month. 12 months in a year. period per year is 12.
*principal is $1500
*we are trying to calculate the final balance after 6 years.
*interest rate is 3.25% which is 0.0325

```{r}

principal <- 1500
years <- 6
interest_rate < 0.0324
period_per_year <- 12 

for(period in 1:(period_per_year*years)){
  principal <- principal + (principal * interest_rate/period_per_year)
}

print(principal)

```
Question 11:Create a numeric vector length 20 and then write a code to calculate the sum of every third element of the vector you have created.
To get the every 3rd element i can use [... %%3]

```{r}

num_vector <- c(1:20)
sum(num_vector[num_vector %% 3])



```

Question 12: Use loop to calculate (could not add special characters) for the value x=2 
Basically the question: get sum of x^1 + x ^2 + x^3 .... up to x^10 for value x being 2

```{r}

x <-2 
sum <- 0

for (i in 1:10){
  sum <- sum + x ^ i
}
print(sum)


```

Question 13: Use a while loop to accomplish the same task as in question 12.

```{r}

x <-2
sum <-2

while(i <=10){
  sum <- sum + x ^ i
  i <- i+1
}
print(sum)

```

Question 14: Solve the problem freom the two previous exercises without using a loop. 
```{r}
x <- 2
sum(x^(1:10))
```




